---
output: github_document
  html_document:lab3
    df_print: paged
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}

av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)

```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`.

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers.
### Thien Nguyen Tam Statement:
> average number of deaths per Avenger is approximately 0.55

```{R}


library(tidyverse)

glimpse(av)

print(names(av))

deaths <- av %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Death_Time",
    values_to = "Death_Status"
  ) %>%
  mutate(
    Death_Time = parse_number(Death_Time),
    Death = case_when(
      Death_Status == "YES" ~ "yes",
      Death_Status == "NO"  ~ "no",
      TRUE                  ~ NA_character_  
    )
  ) %>%
  select(Name.Alias, Death_Time, Death)


print("Preview of Deaths Data:")
head(deaths)


returns <- av %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Return_Time",
    values_to = "Return_Status"
  ) %>%
  mutate(
    Return_Time = parse_number(Return_Time),
    Return = case_when(
      Return_Status == "YES" ~ "yes",
      Return_Status == "NO"  ~ "no",
      TRUE                   ~ NA_character_  
    )
  ) %>%
  select(Name.Alias, Return_Time, Return)


print("Preview of Returns Data:")
head(returns)



death_time_range <- deaths %>%
  summarise(min_time = min(Death_Time, na.rm = TRUE),
            max_time = max(Death_Time, na.rm = TRUE))


deaths_complete <- deaths %>%
  complete(Name.Alias, Death_Time = seq(death_time_range$min_time, death_time_range$max_time)) %>%
  mutate(
    Death = replace_na(Death, "no") 
  )


print("Completed Deaths Data:")
head(deaths_complete)

average_deaths <- deaths_complete %>%
  mutate(Death_Binary = if_else(Death == "yes", 1, 0)) %>%
  group_by(Name.Alias) %>%
  summarise(Total_Deaths = sum(Death_Binary, na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(Average_Deaths = mean(Total_Deaths, na.rm = TRUE))

print("Average Number of Deaths:")
print(average_deaths)



```

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Thien Nguyen Tam Statement:

>Avengers suffer a total of 89 deaths and experience 57 comebacks throughout their history.

```{R}
library(tidyverse)

# Calculate total deaths
total_deaths <- deaths %>%
  filter(Death == "yes") %>%
  tally() %>%
  pull(n)

# Calculate total comebacks
total_comebacks <- returns %>%
  filter(Return == "yes") %>%
  tally() %>%
  pull(n)

# Display the results
total_deaths
total_comebacks

```

Here, the total deaths confirms that there were indeed 89 total deaths and 57 comebacks. This was done by filtering out deaths = "yes" and return = "yes".

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Thien Nguyen Tam Statement:

>The analysis confirms that Iron Man is the Avenger with the most recorded deaths, accumulating a total of 7 deaths


```{r}

library(tidyverse)

# Calculate total deaths per Avenger
death_counts <- deaths_complete %>%
  filter(Death == "yes") %>%
  group_by(Name.Alias) %>%
  summarise(Total_Deaths = n()) %>%
  arrange(desc(Total_Deaths))

# Identify the Avenger with the most deaths
top_death_avenger <- death_counts %>%
  slice(1)

print(top_death_avenger)
```

### Answer

The data analysis reveals that Iron Man has the highest number of recorded deaths among the Avengers, with a total of 7 deaths.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
